---
title: "Figures in for loop"
author: "EE Holmes"
output:
  pdf_document: default
  html_document: default
---


With base R plots, you can just put them in a for loop.

```{r}
library(ggplot2)
for( i in 1:2 ) {
  df <- data.frame(t=1:100, y=rnorm(100))
  plot(df$t, df$y)
  cat('\n\n')
  }
```

With **ggplot2** you need to assign to an object and print that.

```{r}
library(ggplot2)
for( i in 1:2 ) {
  df <- data.frame(t=1:100, y=rnorm(100))
  p <- ggplot(df, aes(t, y)) + geom_point()
  print(p)
  cat('\n\n')
  }
```

To have figure captions that change, you need to create all your figure captions and give that to `fig.cap` as a vector.


```{r fig.cap = paste("Figure", 1:2)}
library(ggplot2)
for( i in 1:2 ) {
  df <- data.frame(t=1:100, y=rnorm(100))
  p <- ggplot(df, aes(t, y)) + geom_point()
  print(p)
  cat('\n\n')
  }
```

You can have figures appear side by side, though using figure captions will break this. You can `cat()` out a legend but you'll need to set `results='asis'`.

```{r echo=FALSE, fig.hold=TRUE, out.width="50%", results='asis'}
library(ggplot2)
for( i in 1:2 ) {
  df <- data.frame(t=1:100, y=rnorm(100))
  plot(df$t, df$y)
  p <- ggplot(df, aes(t, y)) + geom_point()
  print(p)
  cat(paste("Figure", i, "\n\n"))
  }
```