---
title: "rep-res-coho-example"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  github_document:
    toc: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "readme-figs/"
)
```

## General Goals

This is a simple RStudio project that Eric Anderson has put together
as an example/template for students to get a sense of what he expects
them to do in terms of assembling their data for the 2017 EEB 295 course,
"[Case Studies in Reproducible Research](https://eriqande.github.io/rep-res-eeb-2017/)".

The general goals of this project (for our purposes at the moment) are fairly straightforward.
We want to use pedigree information to summarize the ages and distributions of family sizes 
of coho returning to Klamath River hatcheries.  etc. etc.  

## Data

The data for this project are all housed in the `./data` directory.  There are two main types of 
files:

1.  There are three files which are output files from the program [SNPPIT](https://github.com/eriqande/snppit).
These are `snppit_output_ParentageAssignments_2013Juvs.txt`,  `snppit_output_ParentageAssignments_2014Juvs.txt`, and `snppit_output_ParentageAssignments_2015Juvs.txt`.  These are TAB-delimited text files which give the inferred trios
(Father-Mother-Offspring) of hatchery coho salmon in our Shasta River Project.  These files use "---" to denote columns
that have missing data.  These result from analyses made on the
genetic data.  In a real reproducible example, we would have started from the genotype data and actually run the 
SNPPIT analyses reproducibly, as well.  But, for an example, it will be simpler to start from these simple,
intermediate files.

2. There is one data file of extra metadata that should include all the individuals in the snppit output files
(and probably a few extra ones as well.)  The main key between this file and the other ones is the 
NFMS_DNA_ID which is part of the ID in the Kid, Ma, and Pa columns in the `snppit_output*` files.



### Reading Data In

#### SNPPIT files
The SNPPIT files can be read in with `read_tsv()` making note of the missing data "---".
```{r read-one-snppit}
library(tidyverse)

snppit2013 <- read_tsv(file = "data/snppit_output_ParentageAssignments_2013Juvs.txt", 
                       na = "---")
```

Of course, if we wanted to read them all in at once and make a tidy frame of all of them
we would do:
```{r read-all-snppit}
# read all three into a list
years <- 2013:2015
names(years) <- years
snppit <- lapply(years, function(y) {
  read_tsv(file = paste("data/snppit_output_ParentageAssignments_", y, "Juvs.txt", sep = ""),
           na = "---")
})

# note that we can check the column specifications for any one of them
spec(snppit$`2014`)

# and we can put them together into a single data frame with
# a juvie_year column like this
snppit_all <- bind_rows(snppit, .id = "juvie_year") %>%
  mutate(juvie_year = as.integer(juvie_year))
```

#### The Meta Data

This file is actually an interesting example because it has column names with single quotes (who did that?)
and also with "#" symbols.  If we read this with base R's `read.csv()` it will mangle those
names.  The `readr` functions never do that.  There are also clearly some problems, which 
we will get to later
```{r read-meta}
meta <- read_csv("data/shasta-meta-and-genos.csv")
```
To see a tibble of the problems we can do:
```{r meta-probs}
problems(meta)
```