---
title: "Tables"
author: "EE Holmes"
date: "6/12/20202"
output:
  pdf_document: default
  html_document: default
  word_document: default
always_allow_html: yes
---

This is a more complex example of making tables.

```{r setup, warning=FALSE}
library(knitr)
library(kableExtra)
library(xtable)
output.type <- knitr::opts_knit$get('rmarkdown.pandoc.to')
output.type
```

Here I'll use a data frame from from a R data set, but in practice your data frame would come from a function.

Note the first time you make a PDF, tinytex will load any needed packages and it can take a **long** time. You'll see a spinning wheel on the R Markdown tab.

```{r}
a <- mtcars[1:7, 1:6]
```

Default **kable** table.

```{r}
knitr::kable(a)
```

This is the table with **xtable** for LaTeX and **kableExtra** for html.

```{r, results='asis', echo=FALSE}
caption="My table caption"
#Dynamically add stuff to caption
caption <- paste(caption, Sys.Date())
# set up latex alignment
aln <- paste0("rr|c|", paste(rep("c", ncol(a)-2), collapse=""))
if(output.type=="latex"){
  # Google table placement latex to learn the placement options
  print(xtable::xtable(a, align=aln, digits=0, caption=caption), 
      comment=FALSE, include.rownames=FALSE,
      caption.placement="top", table.placement="h", type=output.type)
  
  # Landscape for wide tables. You need table.placement="p" here because you need a new page.
  cat("\\begin{landscape}")
  print(xtable::xtable(a, align=aln, digits=0, caption=caption), 
      comment=FALSE, include.rownames=FALSE,
      caption.placement="top", table.placement="p", type=output.type)
  cat("\\end{landscape}")
  
  # If you just want the tex file
    print(xtable::xtable(a, align=aln, digits=0, caption=caption), 
      comment=FALSE, include.rownames=FALSE,
      caption.placement="top", table.placement="p", type=output.type, file="table.tex")
  }
if(output.type=="html"){
  cat("\n## The table heading\n\n")
  cat(caption)
  knitr::kable(a) %>% 
    kableExtra::kable_styling(full_width = FALSE)
}
if(output.type=="word"){
  cat("\n## The table heading\n\n")
  cat(caption)
  knitr::kable(a)
}
```

