---
title: "Tables using Children and Table Functions"
author: "EE Holmes"
output:
  word_document: default
  html_document: default
  pdf_document: default
always_allow_html: yes
---

```{r include=FALSE}
# include = FALSE means don't show code or any output (like warnings)
# It is good to set up any knitr options that you want at the top of all your Rmds
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
```

This is an example of a Tables Rmd with separate Rmds for the individual tables. `Table1` is a simple table. `Table_Counts` is a table function that is then called to make 5 tables with the same format but different input data. For this example, I am only going to use `kable()` for all output types. See `Tables.Rmd` for examples of using the **kableExtra*` and** and **xtable** packages. 

The example in Table_Counts.Rmd shows you how to include a page break in your Word doc between tables. Also note that I use `format="pandoc"` for the table. That works for Word in my experience.

Also note, best not to use chunk labels in the your Rmd children. It's too easy to get duplicate labels accidentally.

**Make sure these files are in your working directory. You may need to set the working directory to Report_with_Tables.**


```{r child="Table1.Rmd"}
```

```{r child="Table_Counts.Rmd"}
```

<!-- This is how to get a new page in Word. Make sure you are in print view on the word doc -->

```{=openxml}
<w:p><w:r><w:br w:type="page"/></w:r></w:p>
```

```{r child="Table1.Rmd"}
```


